---
title: "Handwritten Digit Recognition using CNN on MNIST"
output: html_notebook
---

This script is the demo code for chapter 1.

Loading data:
```{r}
dat <- read.csv2('train.csv', sep = ',')
```

Explore the dataset:
```{r}
dim(dat)
head(dat[1:6])  # this shows the dataframe is column oriented
unique(unlist(dat[1]))
min(dat[2:785])
max(dat[2:785])
```

Each row of the dataframe represents a grey image (a handwritten digit) consists of a label (what the digit is) and 784 pixels (28 * 28). Each pixel is represented by a number between 0 to 255, which is the brightness of the pixel, for example, 0 means black and 255 means white.

Show the image at the 4th row:
```{r}
sample.4 <- matrix(as.numeric(dat[4, -1]), nrow = 28, byrow = TRUE)
image(sample.4, col = grey.colors(255))
```


Show the image at the 7th row:
```{r}
sample.7 <- matrix(as.numeric(dat[7, -1]), nrow = 28, byrow = TRUE)
image(sample.7, col = grey.colors(255))
```

Give a rotation of 90 degrees clockwise on the image, by fristly reverse elements in each column of the matrix, and then transpose this matrix(why?):
```{r}
rotate <- function(x) {t(apply(x, 2, rev))}
image(rotate(sample.4), col = grey.colors(255))
image(rotate(sample.7), col = grey.colors(255))
```

Transform response variable "label" from integer to factor, to perform classification:
```{r}
is.factor(dat$label)
dat$label <- as.factor(dat$label)
is.factor(dat$label)
```

Summarize the label distribution in counts:
```{r}
summary(dat$label)
```

Combined with proportion:
```{r}
prop <- prop.table(table(dat$label))
cbind(count = table(dat$label), proportion = prop)
```

Classes are balanced.

```{r}
central.block <- c('pixel376', 'pixel377', 'pixel404', 'pixel405')
par(mfrow = c(2, 2))
for (i in 1:9) {
  hist(c(as.matrix(dat[dat$label == i, central.block])),
       main = sprintf('Histogram for digit %d', i),
       xlab = 'Pixel Value')
}
```

上面代码中的`central.block`分析如下：

首先：`pixel376`是列名：
```{r}
names(dat)[378]
names(dat)[379]
names(dat)[406]
names(dat)[407]
head(names(dat))
```

其次，`pixel376` 等4个像素点是比较靠近28 * 28 的图片中心的像素点：
```{r}
376 %/% 28
376 %% 28
404 %/% 28
404 %% 28
```

就是说`pixel376` 位于第13行第12列，而 `dat[dat$label == 1, central.block]` 代表在训练集包含的42000张图片中，
所有被标记为1的图片的靠近中心位置的4个像素点的亮度值。

P.S.: 28 * 28图片的正中心是第14行的第14,15列，第15行的第14,15列，共4个像素，当第一个像素记为`pixel0`时，
第i行第j列的标号是$(i - 1) \times 28 + (j - 1)$，所以这4个像素的标号是 `pixel377`, `pixel378`, `pixel405`, `pixel406`. 所以书中这4个点不是图片正中心，而是正中心向左偏一个像素的4个点。

从直方图可以看到大多数写有数字2,4,7图片的中心是黑色（直方图左侧第一列brightness = 0占比最大），
写有数字3,8图片的中心以白色为主（直方图第右侧brightness = 255占比最大）。

